---
title: "SC-ECO"
subtitle: "Population growth rate (r)"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: word_document
github_document: 
   mathjax: TRUE 
---

```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)

opts_chunk$set(comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE,
               fig.width =8, 
               fig.height=8,
               cache     =TRUE, 
               fig.path  ="../tex/fb-",
               cache.path="cache/unassessed/fb/")

iFig=0
iTab=0
```
```{r, pkgs}  
library(ggplot2)
library(plyr)
library(dplyr)
library(reshape)

library(FLCore)
library(FLBRP)
library(FLasher)
library(ggplotFL)
library(FLife)


library(GGally)

library(mydas)
library(popbio)
```


```{r, data}
load("/home/laurence-kell/Desktop/papers/data-poor/data/fb.RData")
spp=read.csv("/home/laurence-kell/Desktop/rfmo/iccat/sc-eco/2020/data/CODES_Species.csv")

names(spp)[3]="species"
spp$species=as.character(spp$species)
fb$species=as.character(fb$species)
lh=merge(fb,spp,by="species")

names(lh)[c(14:17)]=c("l50","l50min","l50max","a50")
lh=lh[,c("species","genus","family","swimmode","habit","linf","k","t0","a","b","a50","l50","l50min","l50max")]

lh[is.na(lh["l50"]),"l50"]=(lh[is.na(lh["l50"]),"l50min"]+lh[is.na(lh["l50"]),"l50max"])/2
lh=lh[,-(13:14)]

lh[lh$t0>=0,"t0"]=NA

lh=ddply(lh,.(species), with, 
         data.frame(linf=mean(linf,na.rm=TRUE),
                    k   =mean(k,   na.rm=TRUE),
                    t0  =mean(t0,  na.rm=TRUE),
                    a   =mean(a,   na.rm=TRUE),
                    b   =mean(b,   na.rm=TRUE),
                    a50 =mean(a50, na.rm=TRUE),
                    l50 =mean(l50, na.rm=TRUE)))
lh=subset(lh,!is.na(linf)&!is.na(a)&!is.na(b))
spp=lh$species
lh=as(lh[,-1],"FLPar")
lh=lhPar(lh)
dimnames(lh)$iter=spp
```

```{r, save}
save(lh,file="/home/laurence-kell/Desktop/projects/indicators/productivity/data/lh-fb.RData")
```

```{r, scenarios}
design=data.frame(s      =c( 0.9, 0.7,  0.9,  0.9,  0.9,  0.9),
                  sel2   =c( 1.0, 1.0,  5000, 1.0,  1.0,  1.0),
                  sel3   =c(5000, 5000, 5000, 50,   5000, 5000),
                  m2     =c(1,    1,    1,    1,    0.9,  1.1))
```


```{r, run}
r=mdply(data.frame(sp=spp), function(sp) {
  
  pars        =propagate(lh[,sp],dim(design)[1])
  pars["s"]   =design[,"s"]
  pars["sel2"]=design[,"sel2"]
  pars["sel3"]=design[,"sel3"]
  pars["m2"]  =pars["m2"]*design[,"m2"]
  
  eq=try(lhEql(pars))
  
  if ("try-error"%in%is(eq))
    return(NA)
  else
    rtn=try(mdply(data.frame(i=seq(dims(eq)$iter)),function(i) 
      log(lambda(leslie(eq,fbar=c(refpts(eq)["crash","harvest",i]))[,,i,drop=TRUE]))))
  
  if ("try-error"%in%is(rtn)) return(NA) else return(data.frame(r=rtn$V1,design))
 })

r=data.frame(r,model.frame(lh[c("a","b","linf","t0","k","l50"),r$sp]))

save(r,file="/home/laurence-kell/Desktop/projects/indicators/productivity/data/r-fb.RData")
```

```{r, lh.r}
ggpairs(transform(r[,c("r","linf","k","t0","b","l50")],lmatlinf=l50/linf,logr=log(r))[,c(8,2:7)])+
  theme_bw()  
```

**Figure `r iFig=iFig+1; iFig`** Relationship between population growth rate ($r$) and life history parameters


```{r, robustness}
dat=cast(data.frame(r,scenario=1:6)[,c(1,2,14)],sp~scenario,value="r")[,-1]
names(dat)=c("base","h=0.7","flat","dome","low M","high M")

ggpairs(dat)+
  theme_bw()  
```

**Figure `r iFig=iFig+1; iFig`** Correlations between "r" under different assumptions. 


```{r, robustness2}
dat=data.frame(r=r$r,scenario=1:6)
dat=as.data.frame(t(ddply(dat,.(scenario),with,order(r)))[-1,])
names(dat)=c("base","h=0.9","flat","dome","low M","high M")

ggpairs(dat)+
  theme_bw()  
```

**Figure `r iFig=iFig+1; iFig`** 

```{r}
load("/home/laurence-kell/Desktop/projects/indicators/productivity/data/msy.RData")
load("/home/laurence-kell/Desktop/projects/indicators/productivity/data/r-fb.RData")
names(r)[1]="species"

```

\newpage
## Software Versions

* `r version$version.string`
* FLCore:    `r packageVersion('FLCore')`
* FLBRP:     `r packageVersion('FLBRP')`
* FLasher:   `r packageVersion('FLasher')`
* FLife:     `r packageVersion('FLife')`
* mydas:     `r packageVersion('mydas')`
* **Compiled**: `r date()`

## Author information

**Laurence Kell**. laurie@seaplusplus.es

# References {#References}
