---
title: "SC-ECO"
subtitle: "Productivity Analysis"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: word_document
github_document: 
   mathjax: TRUE 
---

## Objective

Compare different ways to estimate productivity for **data poor** stocks and compare with estimates from **data rich** assessments.

## Data Rich Assessments

+ Use data rich assessments from the Myers legacy database to sumarise reference points, i.e. $MSY$, $F_{MSY}$, $B_{MSY}$, and $K$.
+ Use the reference points to estimate the Pella-Tomlinson production functions and hence $r$ and the shape of the production function ($p$).

### Life History characteristics

+ A number of life history characteristics have been used as proxies for productivity, i.e.
   + Maximum Size ($L_{max}$)	<110 cm	110–200	>200 cm
   + von Bertalanffy growth coefficient ($k$)	>0.36	0.27–0.36	<0.27
   + Size at first maturity ($L_{50}$)	<54	54–105	>105 
   + Maximum age ($T_{max}$)	<8	8–14	>14 
   + $L_{50}/L_{max}$	<0.51	0.51–0.55	>0.55
   + Fecundity (Fec)	>2.88	1.03–2.88	<1.03 
   + $r$	>0.48	0.26–0.38	<0.38

+ Use life history parameters from Fish Base, and identify cases where $L_{\infty}$, $k$, $L_{mat}$, and $b$ from the length weight relationship are available and then calculate $r$ under a variety of assumptions and compare to $F_{MSY}$ and $r$ from the data rich case studies.
+ Scenarios corresponding to steepness ($h$), selection pattern (dome shaped, flat) and $M$.

### Productivity Proxies

+ Compare life history parameters and estimates of $r$ with **Data Rich** estimates to develop a robust proxy for productivity.


```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)

opts_chunk$set(comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE,
               fig.width =8, 
               fig.height=8,
               cache     =TRUE, 
               fig.path  ="../tex/p-",
               cache.path="cache/unassessed/p/")

iFig=0
iTab=0
```
```{r, pkgs}  
library(ggplot2)
library(plyr)
library(dplyr)
library(reshape)

library(FLCore)
library(FLBRP)
library(FLasher)
library(ggplotFL)
library(FLife)
library(randtests)

library(mvtnorm)

library(GGally)
library(mydas)
library(popbio)

library(dendextend)

library(doParallel)
library(foreach)
```


```{r, myers}
load("/home/laurence-kell/Desktop/papers/data-poor/data/DBdata.RData")

names(stock)[3]="species"
msy=subset(bioparams,bioid%in%c("SSBmsy-MT","MSY-MT","SSB0-MT"))
msy=cast(msy,stockid+assessid~bioid,value="biovalue")
names(msy)[3:5]=c("msy","K","bmsy")

msy=mutate(msy, msy=as.numeric(as.character( msy)),
                bmsy=as.numeric(as.character(bmsy)),
                hmsy=msy/bmsy)
msy=merge(msy,stock)
msy=subset(msy,!is.na(hmsy))

msy$K=as.numeric(as.character(msy$K))
msy$shape=msy$bmsy/msy$K
msy=adply(msy, 1, function(x) data.frame(p=optimise(function(x,y) (y-(1/(1+x))^(1/x))^2,c(-0.9999,10),y=x$shape)$minimum))
msy=transform(msy,r=(1+p)*(msy/bmsy))

msy=subset(msy,!is.na(K))
msy$species=as.character(msy$species)

save(msy,file="../data/msy.RData")
```


```{r, lh}
load("/home/laurence-kell/Desktop/papers/data-poor/data/fb.RData")
lh=merge(msy,fb,by="species")
lh[is.na(lh["lmat"]),"lmat"]=(lh[is.na(lh["lmat"]),"lmatmin"]+lh[is.na(lh["lmat"]),"lmatmax"])/2
lh$t0[lh$t0>=0]=NA
lh=ddply(lh,.(species), with, data.frame(k   =mean(k,   na.rm=T),
                                         linf=mean(linf,na.rm=T),
                                         t0  =mean(t0,  na.rm=T),
                                         l50 =mean(lmat,na.rm=T),
                                         a50 =mean(amat,na.rm=T),
                                         a   =mean(a,   na.rm=T),
                                         b   =mean(b,   na.rm=T)))
```

```{r}
source('~/Desktop/flr/FLife/R/FLife-lhPar.R')

popdyn2<-function(x) {
  #n<<-n+1;print(n)
  
  if (is.na(x["linf"])) return(NULL) 
  
  x=x[-1]
  x=try(as(x[,names(x)[!is.na(x)]],"FLPar"))
  
  if ("try.error"%in%is(x)) return(NULL)
  
  x=lhPar(x)
  x=try(popdyn(x))
  
  if ("try.error"%in%is(x)) return(NULL) else return(model.frame(x))}

dat=merge(msy,lh)
dat=dat[!duplicated(dat$species),]

n=0
pd=adply(dat[,c("species","linf","k","t0","a","b","l50","a50")],1,popdyn2)

save(pd,file="/home/laurence-kell/Desktop/projects/indicators/productivity/data/pd.RData")
```


```{r}
r=merge(msy[,c("species","K","bmsy","hmsy","shape","p","r")],pd,by="species")
dat=r[,c("r.x","r.y","hmsy","linf","k","l50","shape")]
names(dat)[1:2]=c("r","r.lh")

ggpairs(log(transform(dat[,c("r","hmsy","shape","r.lh","linf","k","l50")],l50linf=l50/linf)))+
  theme_bw()
```

**Figure `r iFig=iFig+1; iFig`** Relationships between $r$, $H_{MSY}$, and the shape of the Pella-Tomlinson production function from data rich assessments, with data poor estimates of $r$ (r.lh) and life history parameters $L_{\nfty}$, $k$ and $L_{50}/L_{\infty}$.
 

```{r, scenarios}
design=data.frame(s      =c( 0.9, 0.7,  0.9,  0.9,  0.9,  0.9),
                  sel2   =c( 1.0, 1.0,  5000, 1.0,  1.0,  1.0),
                  sel3   =c(5000, 5000, 5000, 50,   5000, 5000),
                  m2     =c(1,    1,    1,    1,    0.9,  1.1))
```


```{r, run}
r2=ddply(r[!duplicated(r$species),], .(species), with, {
  pars        =c(linf=linf,k=k,t0=t0,a=a,b=b,l50=l50)
  pars        =lhPar(FLPar(pars))
  pars        =propagate(pars,dim(design)[1])
  pars["s"]   =design[,"s"]
  pars["sel2"]=design[,"sel2"]
  pars["sel3"]=design[,"sel3"]
  pars["m2"]  =pars["m2"]*design[,"m2"]
  
  eq=lhEql(pars)

  #if (!("try-error"%in%is(eq))) 
  data.frame(species=species,r=mdply(data.frame(i=seq(dim(design)[1])),function(i) 
       log(lambda(leslie(eq,fbar=c(refpts(eq)["crash","harvest",i]))[,,i,drop=TRUE])))[,2],
       design)
  })

save(r2,file="/home/laurence-kell/Desktop/projects/indicators/productivity/data/r2.RData")
```


```{r, lh.r}
#load("/home/laurence-kell/Desktop/projects/indicators/productivity/data/r2.RData")
dat=as.data.frame(t(array(r2$r,c(dim(r2)[2],dim(r2)[1]/dim(r2)[2]))))

names(dat)=c("Base","h=0.7","flat","dome","low M","high M")

ggpairs(log(dat))+
  theme_bw()  
```

**Figure `r iFig=iFig+1; iFig`** Relationship between data poor estimates population growth rate ($r$) under the different scenarios.

```{r}
dat=merge(msy,r2,by="species",all=TRUE)
dat=cbind(scen=seq(6),subset(dat,!is.na(r.x)&!is.na(r.y)))

ggplot(dat)+geom_histogram(aes(r.y,fill=hmsy<0.2))+
  facet_grid(scen~.)+
  xlab("r")+ylab("Frequency")
```

**Figure `r iFig=iFig+1; iFig`** Estimates of $r$ for data poor scenarios, colours compare with $H_{MSY}$.


\newpage
## Software Versions

* `r version$version.string`
* FLCore:    `r packageVersion('FLCore')`
* FLBRP:     `r packageVersion('FLBRP')`
* FLasher:   `r packageVersion('FLasher')`
* FLife:     `r packageVersion('FLife')`
* mydas:     `r packageVersion('mydas')`
* **Compiled**: `r date()`

## Author information

**Laurence Kell**. laurie@seaplusplus.es

# References {#References}
  