---
title: "SC-ECO"
subtitle: "Population growth rate (r)"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: word_document
github_document: 
   mathjax: TRUE 
---

```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)

opts_chunk$set(comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE,
               fig.width =8, 
               fig.height=8,
               cache     =TRUE, 
               fig.path  ="../tex/ram-",
               cache.path="cache/unassessed/ram/")

iFig=0
iTab=0
```
```{r, pkgs}  
library(ggplot2)
library(plyr)
library(dplyr)
library(reshape)

library(FLCore)
library(FLBRP)
library(FLasher)
library(ggplotFL)
library(FLife)
library(randtests)

library(mvtnorm)

library(GGally)
library(mydas)
library(popbio)

library(dendextend)

library(doParallel)
library(foreach)
```


```{r, paper-om, fig.height=6, eval=FALSE}
sim<-function(x,niters=500,se=0.1){
  
  mn=aaply(x,1,mean, na.rm=TRUE)
  sd=aaply(x,1,var,  na.rm=TRUE)^0.5
  n =aaply(x,1,function(x) sum(!is.na(x)))
  se=sd/n^0.5
  
  if (any(is.na(se))) se[is.na(se)]=se
  
  y=data.frame(mn=mn,se=se)
  y=mdply(y,function(mn,se) rnorm(niters,mn,se))[,-(1:2)]
  
  res=FLPar(array(unlist(c(y)),c(dim(x)[1],niters)))
  
  dimnames(res)$params=names(mn)
  
  res}

cor=cor(model.frame(lh)[,c("linf","k","l50","t0","a","b","l50linf")],
        use="pairwise.complete.obs")
lhs=daply(lh,.(species), with, 
             FLPar(linf=linf,k=k,t0=t0,l50=l50,a=a,b=b,bg=b,l50linf=l50/linf))
```

```{r}
load("/home/laurence-kell/Desktop/papers/data-poor/data/DBdata.RData")
load("/home/laurence-kell/Desktop/papers/data-poor/data/fb.RData")
```


```{r, msy}
names(stock)[3]="species"
msy=subset(bioparams,bioid%in%c("SSBmsy-MT","MSY-MT","SSB0-MT"))
msy=cast(msy,stockid+assessid~bioid,value="biovalue")
names(msy)[3:5]=c("msy","K","bmsy")

msy=mutate(msy, msy=as.numeric(as.character( msy)),
                bmsy=as.numeric(as.character(bmsy)),
                hmsy=msy/bmsy)
msy=merge(msy,stock)
msy=subset(msy,!is.na(hmsy))

msy$K=as.numeric(as.character(msy$K))
msy$shape=msy$bmsy/msy$K
msy=adply(msy, 1, function(x) data.frame(p=optimise(function(x,y) (y-(1/(1+x))^(1/x))^2,c(-0.9999,10),y=x$shape)$minimum))
msy=transform(msy,r=(1+p)*(msy/bmsy))

msy=subset(msy,!is.na(K))
msy$species=as.character(msy$species)

save(msy,file="../data/msy.RData")
```

```{r, lh}
ref=merge(msy,fb,by="species")
ref[is.na(ref["lmat"]),"lmat"]=(ref[is.na(ref["lmat"]),"lmatmin"]+ref[is.na(ref["lmat"]),"lmatmax"])/2

lh=ddply(ref,.(species), with, data.frame(k   =mean(k,   na.rm=T),
                                          linf=mean(linf,na.rm=T),
                                          t0  =mean(t0,  na.rm=T),
                                          l50 =mean(lmat,na.rm=T),
                                          a50 =mean(amat,na.rm=T),
                                          a   =mean(a,   na.rm=T),
                                          b   =mean(b,   na.rm=T)))

lh=subset(lh,!is.na(a)&t0<0&!is.na(l50))
```


```{r}
source('~/Desktop/flr/FLife/R/FLife-lhPar.R', echo=TRUE)

popdyn2<-function(x) {
  n<<-n+1
  
  if (is.na(x["linf"])) return(NULL) 
  
  x=x[-1]
  x=try(as(x[,names(x)[!is.na(x)]],"FLPar"))
  
  if ("try.error"%in%is(x)) return(NULL)
  
  x=lhPar(x)
  x=try(popdyn(x))
  
  if ("try.error"%in%is(x)) return(NULL) else return(model.frame(x))}

n=0
pd=adply(lh[,c("species","linf","k","t0","a","b","l50")],1,popdyn2)

save(pd,file="../data/pd.RData")
```

```{r}
r=merge(msy[,c("species","K","bmsy","hmsy","shape","p","r")],pd,by="species")
dat=r[,c("r.x","r.y","hmsy","linf","k","shape")]
names(dat)[1:2]=c("r","r.lh")

ggpairs(dat[,c("r","hmsy","shape","r.lh","linf","k")])+
  theme_bw()
```

**Figure `r iFig=iFig+1; iFig`** Relationship between population growth rate ($r.lh$) based on life history parameters alone and values from data rich assessments.
 

```{r}
r2=transform(r, r.lh =order(r.y),
                r    =order(r.x),
                hmsy =order(hmsy),
                linf =order(linf),
                k    =order(k),
                shape=order(shape))
ggpairs(r2[,c("r","hmsy","shape","r.lh","linf","k")])+
  theme_bw()
```

**Figure `r iFig=iFig+1; iFig`** Relationship between population growth rate ($r$) and life history parameters and assessment outputs.


```{r, eval=FALSE}
ts=subset(timeseries,assessid%in%unique(msy$assessid)&!is.na(tsvalue))
ts=cast(ts,assessid+tsyear~tsid,value="tsvalue")
names(ts)[2:4]=c("year","ssb","catch")

pe=merge(msy[,c("species","assessid","r","K","p")],ts,by="assessid")
pe=subset(pe,!is.na(K))
pe=ddply(pe,.(assessid,species), with, {b=c((ssb[-1]+ssb[-length(ssb)])/2,NA); 
                                 data.frame(year =year,
                                            ssb  =ssb,
                                            K    =K,
                                            catch=catch,
                                            sp   =r/p*b*(1-(b/K)^p),
                                            pe   =c(ssb[-1],NA)-c(ssb[-length(ssb)],NA)+catch+r/p*b*(1-(b/K)^p))})

save(pe,file="../data/pe.RData")
```


\newpage
## Software Versions

* `r version$version.string`
* FLCore:    `r packageVersion('FLCore')`
* FLBRP:     `r packageVersion('FLBRP')`
* FLasher:   `r packageVersion('FLasher')`
* FLife:     `r packageVersion('FLife')`
* mydas:     `r packageVersion('mydas')`
* **Compiled**: `r date()`

## Author information

**Laurence Kell**. laurie@seaplusplus.es

# References {#References}
  