---
title: "Data poor validation" 
subtitle: "JABBA MP: ICCAT SS & JABBA OM"
author: "Laurence Kell"
output: rmarkdown:::pdf_document
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
tags: FLPKG FLR
license: Creative Commons Attribution-ShareAlike 4.0 International
---
  
```{r knitr_init, echo=FALSE, results="hide"}
library(knitr)
## Global options
opts_chunk$set(echo    =FALSE,
               eval    =TRUE,
               cache   =TRUE,
               cache.path="../cache/",
               prompt  =FALSE,
               comment =NA,
               message =FALSE,
               tidy    =FALSE,
               warning =FALSE,
               fig.height=6,
               fig.width =6,
               fig.path  ="../tex/jabba",
               dev       ="png")

options(digits=3)

iFig=0
```

```{r, pkgs, echo=FALSE, message=FALSE}
library(reshape)
library(plyr)
library(dplyr)
library(ggplot2)

library(sraplus)
```

```{r, data, fig.height=6, fig.width=6}
load("/home/laurence-kell/Desktop/rfmo/iccat/inputs/jabba/jabbaiccat.RData")

jabbaiccat=jabbaiccat[sort(names(jabbaiccat))]
names(jabbaiccat)[3]="swo"

spp=c("Thunnus obesus",
      "Kajikia albida",
      "Xiphias gladius",
      "Makaira nigricans",
      "Thunnus albacares")
names(spp)=names(jabbaiccat)
jb=ldply(jabbaiccat,function(x)
       data.frame(year=x$timeseries$year,b=x$timeseries$ssb,b_div_bmsy=x$timeseries$ssb/x$refpts$bmsy[1]))
names(jb)[1]="stk"
```


```{r, bd, eval=!FALSE}
bds=mlply(data.frame(stk=names(spp)), function(stk){  
  
  x=jabbaiccat[[stk]]
  
  cdp<-try(sraplus::format_driors(
    use_heuristics=!TRUE,
    taxa          =spp[stk],
    
    index      = x$timeseries$ssb,
    index_year = x$timeseries$year,
    catch      = x$timeseries$catch,
    years      = x$timeseries$year
    ))
  
  fit<-try(sraplus::fit_sraplus(
    driors = cdp,      
    engine = "tmb",
    model  = "sraplus_tmb",
    estimate_qslope = FALSE,
    estimate_proc_error = TRUE))

  if ("try-error"%in%is(fit)) return(NULL)
  
  fit})

save(bds,file="../results/jabbaBDResults.RData")
```


```{r, bd10, eval=!FALSE}
bd10=mlply(data.frame(stk=names(spp)), function(stk){
  
  x=jabbaiccat[[stk]]  
  x$timeseries=subset(x$timeseries,year>=max(year)-9)
  
  cdp<-try(sraplus::format_driors(
    use_heuristics=!TRUE,
    taxa          =spp[stk],
    
    index      = x$timeseries$ssb,
    index_year = x$timeseries$year,
    catch      = x$timeseries$catch,
    years      = x$timeseries$year
    ))
  
  fit<-try(sraplus::fit_sraplus(
    driors = cdp,      
    engine = "tmb",
    model  = "sraplus_tmb",
    estimate_qslope = FALSE,
    estimate_proc_error = TRUE))

  if ("try-error"%in%is(fit)) return(NULL)
  
  fit})


save(bd10,file="../results/jabbaBD10Results.RData")
```

```{r, bd4, eval=!FALSE}
bd4=mlply(data.frame(stk=names(spp)), function(stk){
  
  x=jabbaiccat[[stk]]

  x$timeseries=subset(x$timeseries,year>=1960)

  set.seed(4567)  
  
  cdp<-try(sraplus::format_driors(
    use_heuristics=!TRUE,
    taxa          =spp[stk],
    
    index      = x$timeseries$ssb*rev(rlnorm(length(x$timeseries$ssb),0,0.4)),
    index_year = x$timeseries$year,
    catch      = x$timeseries$catch,
    years      = x$timeseries$year
    ))
  
  fit<-try(sraplus::fit_sraplus(
    driors = cdp,      
    engine = "tmb",
    model  = "sraplus_tmb",
    estimate_qslope = FALSE,
    estimate_proc_error = TRUE))

  if ("try-error"%in%is(fit)) return(NULL)
  
  fit})

save(bd4,file="../results/jabbaBD4Results.RData")
```

```{r, bd410, eval=!FALSE}
bd410=mlply(data.frame(stk=names(spp)), function(stk){
  
  x=jabbaiccat[[stk]]
  x$timeseries=subset(x$timeseries,year>=max(year)-9)
  
  set.seed(4567)  
  
  cdp<-try(sraplus::format_driors(
    use_heuristics=!TRUE,
    taxa          =spp[stk],
    
    index      = x$timeseries$ssb*rev(rlnorm(length(x$timeseries$ssb),0,0.4)),
    index_year = x$timeseries$year,
    catch      = x$timeseries$catch,
    years      = x$timeseries$year
    ))
  
  fit<-try(sraplus::fit_sraplus(
    driors = cdp,      
    engine = "tmb",
    model  = "sraplus_tmb",
    estimate_qslope = FALSE,
    estimate_proc_error = TRUE))

  if ("try-error"%in%is(fit)) return(NULL)
  
  fit})  

save(bd410,file="../results/jabbaBD410Results.RData")
```

```{r, sra, eval=!FALSE}
sras=mlply(data.frame(stk=names(spp)), function(stk){
  
  x=jabbaiccat[[stk]]  

  cdp<-try(sraplus::format_driors(
    use_heuristics=TRUE,
    taxa          =spp[stk],
    
    catch      = x$timeseries$catch,
    years      = x$timeseries$year
  ))
  
  fit<-try(sraplus::fit_sraplus(
    driors = cdp,
    engine = "sir",
    adapt_delta = 0.9,
    max_treedepth = 10,
    n_keep = 4000,
    chains = 1,
    cores = 1,
    estimate_qslope = FALSE,
    estimate_proc_error = TRUE))
  
  if ("try-error"%in%is(fit)) return(NULL)
  
  fit})

save(sras,file="../results/jabbaSRAResults.RData")
```

```{r, sra10, eval=!FALSE}
sras10=mlply(data.frame(stk=names(spp)), function(stk){
  
  x=jabbaiccat[[stk]]  
  x$timeseries=subset(x$timeseries,year>=max(year)-9)
  
  cdp<-try(sraplus::format_driors(
    use_heuristics=TRUE,
    taxa          =spp[stk],
    
    catch      = x$timeseries$catch,
    years      = x$timeseries$year
  ))
  
  fit<-try(sraplus::fit_sraplus(
    driors = cdp,
    engine = "sir",
    adapt_delta = 0.9,
    max_treedepth = 10,
    n_keep = 4000,
    chains = 1,
    cores = 1,
    estimate_qslope = FALSE,
    estimate_proc_error = TRUE))
  
  if ("try-error"%in%is(fit)) return(NULL)
  
  fit})

save(sras10,file="../results/jabbaSRA10Results.RData")
```


```{r, srap, eval=!FALSE}
srap=mlply(data.frame(stk=names(spp)), function(stk){
  
  x=jabbaiccat[[stk]]
  r=x$pfunc[,"r"]
  k=x$pfunc[,"k"]
  p=x$pfunc[,"p"]
  
  cdp<-try(sraplus::format_driors(
    use_heuristics=!TRUE,
    taxa          =spp[stk],
    catch         =x$timeseries$catch,
    years         =x$timeseries$year,
    
    shape_prior       =p+1,
    growth_rate_prior =r,
    k_prior           =k,
    initial_state     =0.75, #x$timeseries$ssb[1]/x$refpts[1,"k"],
    terminal_state    =0.2, #x$timeseries$ssb[length(x$timeseries$ssb)]/x$refpts[1,"k"],

    k_prior_cv          =0.3,
    growth_rate_prior_cv=0.3,
    shape_prior_cv      =0.3,
    initial_state_cv    =0.3,
    terminal_state_cv   =0.5
  ))
  
  fit<-try(sraplus::fit_sraplus(
    driors = cdp,
    engine = "sir",
    adapt_delta = 0.9,
    max_treedepth = 10,
    n_keep = 4000,
    chains = 1,
    cores = 1,
    estimate_qslope = FALSE,
    estimate_proc_error = TRUE))
  
  if ("try-error"%in%is(fit)) return(NULL)
  
  fit})

save(srap,file="../results/jabbaSRPResults.RData")
```

```{r, srap10, eval=!FALSE}
srap10=mlply(data.frame(stk=names(spp)), function(stk){

  x=jabbaiccat[[stk]]
  x$timeseries=subset(x$timeseries,year>=max(year)-9)
  r=x$pfunc[,"r"]
  k=x$pfunc[,"k"]
  p=x$pfunc[,"p"]
  
  cdp<-try(sraplus::format_driors(
    use_heuristics=!TRUE,
    taxa          =spp[stk],
    catch         =x$timeseries$catch,
    years         =x$timeseries$year,
    
    shape_prior       =p+1,
    growth_rate_prior =r,
    k_prior           =k,
    initial_state     =0.75, #x$timeseries$ssb[1]/x$refpts[1,"k"],
    terminal_state    =0.2, #x$timeseries$ssb[length(x$timeseries$ssb)]/x$refpts[1,"k"],

    k_prior_cv          =0.3,
    growth_rate_prior_cv=0.3,
    shape_prior_cv      =0.3,
    initial_state_cv    =0.3,
    terminal_state_cv   =0.5
  ))
  
  fit<-try(sraplus::fit_sraplus(
    driors = cdp,
    engine = "sir",
    adapt_delta = 0.9,
    max_treedepth = 10,
    n_keep = 4000,
    chains = 1,
    cores = 1,
    estimate_qslope = FALSE,
    estimate_proc_error = TRUE))
  
  if ("try-error"%in%is(fit)) return(NULL)
  
  fit})  

save(srap10,file="../results/jabbaSRP10Results.RData")
```

```{r, eval=!FALSE}
library(fishmethods)
source('~/Desktop/flr/fishmethods/R/catchmsy.R')

cmsy=mlply(data.frame(stk=names(spp)), function(stk){

  x=jabbaiccat[[stk]]
  r=x$pfunc[,"r"]
  k=x$pfunc[,"k"]
  p=x$pfunc[,"p"]
  
  catch =x$timeseries$catch
  year  =x$timeseries$year

  fit<-catchmsy(
      year   =year,
      catch  =catch,
      catchCV=NULL,
      catargs=list(dist="none",low=0,up=Inf,unit="MT"),
      l0=list(low=0.1, #max(1,min(x$timeseries$ssb[1]/x$refpts[1,"k"],1),
              up =0.8, #min(x$timeseries$ssb[1]/x$refpts[1,"k"],1),
              step=0),
      lt=list(low  =0.1, #rev(x$timeseries$ssb)[1]/x$refpts[1,"k"]*0.5,
              up   =0.8, #rev(x$timeseries$ssb)[1]/x$refpts[1,"k"]*2.0,
              refyr=max(x$timeseries$year)),sigv=0,
      k=list(dist="unif",low=k*0.1,up=k*10),
      r=list(dist="unif",low=r*0.1,up=r*10),
      M=list(dist="unif",mean=0.2,up=0.2,mean=0.00,sd=0.00),
      nsims=30000,
      graphs=NULL)

  fit})

save(cmsy,file="../results/jabbaCMSYResults.RData")
```

```{r, eval=!FALSE}
cmsy10=mlply(data.frame(stk=names(spp)), function(stk){

  x=jabbaiccat[[stk]]
  x$timeseries=subset(x$timeseries,year>=max(year)-9)
  r=x$pfunc[,"r"]
  k=x$pfunc[,"k"]
  p=x$pfunc[,"p"]
  
  catch =x$timeseries$catch
  year  =x$timeseries$year

  fit<-catchmsy(
      year   =year,
      catch  =catch,
      catchCV=NULL,
      catargs=list(dist="none",low=0,up=Inf,unit="MT"),
      l0=list(low=0.1, #max(1,min(x$timeseries$ssb[1]/x$refpts[1,"k"],1),
              up =0.8, #min(x$timeseries$ssb[1]/x$refpts[1,"k"],1),
              step=0),
      lt=list(low  =0.1, #rev(x$timeseries$ssb)[1]/x$refpts[1,"k"]*0.5,
              up   =0.8, #rev(x$timeseries$ssb)[1]/x$refpts[1,"k"]*2.0,
              refyr=max(x$timeseries$year)),sigv=0,
      k=list(dist="unif",low=k*0.1,up=k*10),
      r=list(dist="unif",low=r*0.1,up=r*10),
      M=list(dist="unif",mean=0.2,up=0.2,mean=0.00,sd=0.00),
      nsims=30000,
      graphs=NULL)

  fit})

save(cmsy10,file="../results/jabbaCMSY10Results.RData")
```

```{r}
load("../results/jabbaBDResults.RData")  
load("../results/jabbaBD4Results.RData")  
load("../results/jabbaSRAResults.RData")
load("../results/jabbaSRPResults.RData")
load("../results/jabbaCMSYResults.RData")  

bd=ldply(bds,function(x){
  if (is.null(x)) return(NULL)
  subset(x$results,variable%in%c("b_div_bmsy","b"))})

bd4=ldply(bd4,function(x){
  if (is.null(x)) return(NULL)
  subset(x$results,variable%in%c("b_div_bmsy","b"))})

sra=ldply(sras,function(x){
  if (is.null(x)) return(NULL)
  subset(x$results,variable%in%c("b_div_bmsy","b"))})

srp=ldply(srap,function(x){
  if (is.null(x)) return(NULL)
  subset(x$results,variable%in%c("b_div_bmsy","b"))})

cmsy.bmsy=ldply(cmsy,function(x){
  dt=cbind(x$timeseries,bmsy=rep(x$Values$Bmsy,rep=length(unique(x$timeseries$year))))
  ddply(dt, .(flag,year), with, quantile(value/bmsy,probs=c(0.975,0.945,0.5,0.045,0.025)))
  })
names(cmsy.bmsy)[c(5:7)]=c("upper","mean","lower")

cmsy.b=ldply(cmsy,function(x){
  dt=cbind(x$timeseries,bmsy=rep(x$Values$Bmsy,rep=length(unique(x$timeseries$year))))
  ddply(dt, .(flag,year), with, quantile(value,probs=c(0.975,0.945,0.5,0.045,0.025)))
  })
names(cmsy.b)[c(5:7)]=c("upper","mean","lower")

dat=rbind.fill(cbind(Model="Catch & Index",        bd),
               cbind(Model="Catch & Index Error",  bd4),
               cbind(Model="Catch & Heuristics",   sra),
               cbind(Model="Catch & Priors",       srp),
               cbind(Model="Catch-MSY",  subset(cmsy.b,   flag==1),variable="b"),
               cbind(Model="Catch-MSY",  subset(cmsy.bmsy,flag==1),variable="b_div_bmsy"))[c(1:4,6:8)]

save(dat,file="../results/jbAll.RData")
```


```{r}
ggplot( subset(dat,variable=="b"))+   
  geom_ribbon(aes(year,ymax=upper,ymin=lower),fill="red",alpha=0.25)+
  geom_line(aes(year,mean),col="red")+
  geom_line(aes(year,b),data=jb)+
  facet_grid(stk~Model,scale="free")+
  xlab("Biomass")+ylab("Year")+
  theme_bw()  
```

**Figure `r iFig=iFig+1; iFig`** Absolute trends in abundance for all stocks and methods (black line are the true values).


```{r}
ggplot( subset(dat,variable!="b"))+ 
  geom_ribbon(aes(year,ymax=upper,ymin=lower),fill="red",alpha=0.25)+
  geom_line(aes(year,mean),col="red")+
  geom_line(aes(year,b_div_bmsy),data=jb)+
  facet_grid(stk~Model,scale="free")+
  xlab("Biomass")+ylab("Year")+
  theme_bw()  
```

**Figure `r iFig=iFig+1; iFig`** Trends in abundance relative to $B/B_{MSY}$ for all stocks and methods (black line are the true values).

```{r}
load("../results/jbAll.RData")

ggplot(ddply(subset(dat,variable!="b"),.(stk,variable), subset, year>max(year)-9))+ 
  geom_hline(aes(yintercept=1))+
  geom_ribbon(aes(year,ymax=upper,ymin=lower),fill="red",alpha=0.25)+
  geom_line(aes(year,mean),col="red")+
  geom_line(aes(year,b_div_bmsy),data=ddply(jb,.(stk), subset, year>max(year)-9))+
  facet_grid(stk~Model)+
  xlab(expression(B:B[MSY]))+ylab("Year")+
  theme_bw()  
```

**Figure `r iFig=iFig+1; iFig`** Trends relative to $B/B_{MSY}$ in last 10 years using all the data.


## 10 Years

```{r}
load("../results/jabbaBD10Results.RData")
load("../results/jabbaBD410Results.RData")  
load("../results/jabbaSRA10Results.RData")
load("../results/jabbaSRP10Results.RData")
load("../results/jabbaCMSY10Results.RData")  
  
bd=ldply(bd10,function(x){
  if (is.null(x)) return(NULL)
  subset(x$results,variable%in%c("b_div_bmsy","b"))})

bd4=ldply(bd410,function(x){
  if (is.null(x)) return(NULL)
  subset(x$results,variable%in%c("b_div_bmsy","b"))})

sra=ldply(sras10,function(x){
  if (is.null(x)) return(NULL)
  subset(x$results,variable%in%c("b_div_bmsy","b"))})

srp=ldply(srap10,function(x){
  if (is.null(x)) return(NULL)
  subset(x$results,variable%in%c("b_div_bmsy","b"))})

cmsy.bmsy=ldply(cmsy10,function(x){
  dt=cbind(x$timeseries,bmsy=rep(x$Values$Bmsy,rep=length(unique(x$timeseries$year))))
  ddply(dt, .(flag,year), with, quantile(value/bmsy,probs=c(0.975,0.945,0.5,0.045,0.025)))
  })
names(cmsy.bmsy)[c(5:7)]=c("upper","mean","lower")

cmsy.b=ldply(cmsy10,function(x){
  dt=cbind(x$timeseries,bmsy=rep(x$Values$Bmsy,rep=length(unique(x$timeseries$year))))
  ddply(dt, .(flag,year), with, quantile(value,probs=c(0.975,0.945,0.5,0.045,0.025)))
  })
names(cmsy.b)[c(5:7)]=c("upper","mean","lower")

dat10=rbind.fill(cbind(Model="Catch & Index",        bd),
                 cbind(Model="Catch & Index Error",  bd4),
                 cbind(Model="Catch & Heuristics",   sra),
                 cbind(Model="Catch & Priors",       srp),
                 cbind(Model="Catch-MSY",   subset(cmsy.b,   flag==1),variable="b"),
                 cbind(Model="Catch-MSY",   subset(cmsy.bmsy,flag==1),variable="b_div_bmsy"))[c(1:4,6:8)]

save(dat10,file="../results/jb10.RData")
```


```{r}
ggplot(ddply(subset(dat10,variable!="b"),.(stk,variable), subset, year>max(year)-9))+   
  geom_hline(aes(yintercept=1))+
  geom_ribbon(aes(year,ymax=upper,ymin=lower),fill="blue",alpha=0.25)+
  geom_line(aes(year,mean),col="blue")+
  geom_line(aes(year,b_div_bmsy),data=ddply(jb,.(stk), subset, year>max(year)-9))+
  facet_grid(stk~Model)+
  xlab(expression(B:B[MSY]))+ylab("Year")+  
  theme_bw()
```

**Figure `r iFig=iFig+1; iFig`** Trends relative to $B/B_{MSY}$ in last 10 years using truncated data.


```{r}
load("../results/jb10.RData")  
  
dt10=ddply(subset(dat10,variable!="b"),.(stk,variable), subset, year>max(year)-9)

ggplot(ddply(subset(dat,variable!="b"),.(stk,variable), subset, year>max(year)-9))+ 
  geom_hline(aes(yintercept=1))+
  geom_ribbon(aes(year,ymax=upper,ymin=lower),fill="red",alpha=0.25)+
  geom_line(aes(year,mean),col="red")+
  geom_ribbon(aes(year,ymax=upper,ymin=lower),fill="blue",alpha=0.25,data=dt10)+
  geom_line(aes(year,mean),col="blue",data=dt10)+
  geom_line(aes(year,b_div_bmsy),data=ddply(jb,.(stk), subset, year>max(year)-9))+
  facet_grid(stk~Model)+
  xlab(expression(B:B[MSY]))+ylab("Year")+
  theme_bw()
```

**Figure `r iFig=iFig+1; iFig`** Trends relative to $B/B_{MSY}$ in last 10 years compared for all and truncated data.


## Indicators
```{r}
twothree<-function(x,n1=2,n2=3){
    
  if (length(x)<n1+n2) return(NULL)
  
  mean(x[length(x)-seq(n1)+1])/mean(x[length(x)-(n1+n2):n2+1])}

trend<-function(x,n=10){
    
    if (length(x)<n) return(NA)
    
    y=x[length(x)-(n:1)+1]
    y=y/mean(y)
    
    if (any(is.na(x))) return(NA)
    
    lm(y~x,data=data.frame(x=seq(n),y=y))$coef[2]}
```


```{r}
load("../data/om.RData")  
load("../results/jb10.RData")
res<-cbind("Yrs"="10",dat10)
load("../results/jbAll.RData")
res<-rbind(res,cbind("Yrs"="All",dat))
res=res[,c("Yrs","Model","stk","mean","year")]
names(res)[4]="bbmsy"
om=cbind(Model="OM",subset(om,OM=="Jabba")[,c("stk","bbmsy","year")])
res=rbind(res,cbind(Yrs=10,om),cbind(Yrs="All",om))
```

```{r}
tab=ddply(res, .(stk,Model,Yrs), with, rbind(c(variable="t23",value=twothree(bbmsy,5)),
                                             c(variable="t5", value=trend(   bbmsy,5))))
tab$value=as.numeric(as.character(tab$value))
  
signif(cast(tab,variable+Model~stk~Yrs),2)
```


```{r}
write.csv(tab,file="../tex/tab.cvs")  
```